---
title: "README"
output:
  md_document:
    variant: gfm
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 7,
                      fig.height = 7,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "man/figures/"
)

```


# GlobalRiverMethane (GRiMe)

A repo to analyze global dataset of methane concentrations and fluxes in flowing waters

### Setup

This repo only contains scripts and example data. The data, analyses, figures, etc. are saved on your local dropbox folder. To use the functions and scripts in this package, you will create an R object with the path to your local dropbox directory. You should only need to do this once, and all functions will access and create files within that folder.  

```{r}
path_to_dropbox <- "Your/local/dropbox/folder"
```
### Example

```{r}
# Example code
```

### More to be added

```{r}
# Example code
```

