---
title: "README"
output:
  md_document:
    variant: gfm
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 7,
                      fig.height = 7,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "man/figures/"
)

```


# GlobalRiverMethane (GRiMe) 

A repo to enable Emily's methane problem. Bonus features include functions and scripts to analyze a global dataset of methane in flowing waters. 

### Setup

This repo only contains scripts and example data. The data, analyses, figures, etc. are saved on your local dropbox folder. To use the functions and scripts in this package, you will create an R object with the path to your local dropbox directory. You should only need to do this once, and all functions will access and create files within that folder.  

```{r}
#Replace with your own dropbox folder location
path_to_dropbox <- "C:/Users/emsta/Dropbox/MethDB2.0" #Emily's folder
path_to_dropbox <- "C:/Users/lloken/OneDrive - DOI/GlobalRiverMethane" #Luke's USGS computer

```
There is a .rda file in the dropbox folder that contains several data tables from the MethDB excel file. Some of column names have been changed to remove blanks. These include `conc_df`, `flux_df`, `sites_df`, `papers_df`, and `gis_df`. Both the concentrations and flux tables have been converted to a common unit (uM and mmol m-2 d-1) for all variables (CH4, CO2, N2O, nutrients)

```{r}

library(dplyr)
library(ggplot2)
library(grid)

#Load custom ggplot functions
source("R/ggplot2_utils.R")

# load formatted and converted tables into your R environment
load(file.path(path_to_dropbox, "db_processingR", 
               "MethDB_tables_converted.rda"))

head(conc_df)

head(flux_df)

head(sites_df)

head(papers_df)

```

### Example figures

Let's plot the distribution of mean CH4 concentration. 

```{r, fig.width = 4, fig.height = 4}

commontheme <- list(
  theme_bw(), 
  scale_x_log10nice(omag = seq(-20, 20, 2)), 
  scale_y_continuous(expand = expansion(mult = c(0, .05)))
)

CH4mean_hist_fig <- ggplot(conc_df) +
  geom_histogram(aes(x = CH4mean)) +
  commontheme

print(CH4mean_hist_fig)

# Save image to dropbox or another folder
# ggsave(file.path(path_to_dropbox, "db_processingR",
#                  "Figures", "CH4_meanconc_hist.png"),
#        CH4mean_hist_fig)

```
Let's plot the distribution of mean CH4, CO2, and N2O concentration and fluxes. 

```{r, fig.width = 8, fig.height = 6}

CO2mean_hist_fig <- ggplot(conc_df) +
  geom_histogram(aes(x = CO2mean)) +
  commontheme +
  theme(axis.title.y = element_blank())

# print(CO2mean_hist_fig)

N2Omean_hist_fig <- ggplot(conc_df) +
  geom_histogram(aes(x = N2Omean)) +
  commontheme +
  theme(axis.title.y = element_blank())

# print(N2Omean_hist_fig)

CH4meanflux_hist_fig <- ggplot(flux_df) +
  geom_histogram(aes(x = DiffusiveCH4FluxMean)) +
  commontheme

# print(CH4meanflux_hist_fig)

CO2meanflux_hist_fig <- ggplot(flux_df) +
  geom_histogram(aes(x = CO2FluxMean)) +
  commontheme +
  theme(axis.title.y = element_blank())

# print(CO2meanflux_hist_fig)

N2Omeanflux_hist_fig <- ggplot(flux_df) +
  geom_histogram(aes(x = N2OFluxMean)) +
  commontheme + 
  theme(axis.title.y = element_blank())

# print(N2Omeanflux_hist_fig)

# To save this figure use package "grid" 
# Need to uncomment png() and dev.off()

# png(file.path(path_to_dropbox, "db_processingR",
#                  "Figures", "Mean_gas_concflux.png"),
#     height = 6, width = 8, units = "in", res = 300)

grid.newpage()
plots <- grid.draw(rbind(cbind(ggplotGrob(CH4mean_hist_fig),
                               ggplotGrob(CO2mean_hist_fig), 
                               ggplotGrob(N2Omean_hist_fig)),
                         cbind(ggplotGrob(CH4meanflux_hist_fig),
                               ggplotGrob(CO2meanflux_hist_fig),
                               ggplotGrob(N2Omeanflux_hist_fig)),
                         size = "first"))

# dev.off()

```

